---
title: "Bio-Optical Model"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

```{r}
library(glue)
library(tidyverse)
library(readxl)
library(lubridate)
```

This is a test. 

```{r}
my_aq_site <- "BLWA"
my_aop_yr <- "2017"
my_loc_type <- 'buoy.c0'
```

## Measured Reflectance Spectra

```{r}
spectra_dir <- "/Volumes/hondula/DATA/spectra_buff5m"
spectra_file <- fs::dir_ls(spectra_dir, regexp = glue("{my_aq_site}_{my_aop_yr}"))

spectra_df <- read_csv(spectra_file)
my_spectra_pt <- spectra_df %>% filter(my_cellid)

rrs <- spectra_df %>%
  ggplot(aes(wavelength, reflectance)) +
  geom_line(aes(group = cell)) +
  geom_line(data = my_spectra_pt, col = "green", lwd = 1) +
  theme_bw() +
  coord_cartesian(ylim = c(0, 0.1)) +
  theme(legend.position = "none")
rrs + ggtitle(glue("Measured reflectance at {my_aq_site} {my_loc_type} {my_aop_yr}"))

```

```{r}
my_wavelengths_rs <- my_spectra_pt$wavelength
range(my_wavelengths_rs)
```

> TODO: striping at the lowest measured wavelengths. real? waves?

> TODO: errors for PRLA (2017, 2019, 2020) and PRPO (2017) spectra

> TODO: BLUE 2017 (spectra S2 point in the tree canopy)

> TODO: TOOK 2019 reflectance not published

> TODO: FLNT 2019 reflectance not published

> TODO: curate reflectance at stream sites. look at pics. optically shallow?

For optically shallow waters, would also need bottom depth and bottom albedo,
and diffuse attentuation coefficient ($K$).
See eg. Albert and Mobey 2003 equations

## Flight metadata

Use dates and timing of flight data to get the closest in situ data

Dates of flights from table

```{r}
my_aop_dates <- "../neon-sites/results/sites_join_aop_dates.csv" %>%
  readr::read_csv(col_types = "ccccccccddD") %>%
  dplyr::filter(siteID %in% my_aq_site, year %in% my_aop_yr) %>%
  dplyr::select(domanID, siteID, aop_site_id, flightdate) %>%
  distinct()
my_aop_dates
```

Time of pixel acquisition from L1 radiance data

```{r}
radtimes_dir <- "/Volumes/hondula/DATA/radiance-time/"
radtimes_file <- fs::dir_ls(radtimes_dir, regexp = glue("{my_aq_site}-{my_aop_yr}"))
radtimes_df <- read_csv(radtimes_file) %>% dplyr::filter(!is.na(gps_time))

my_radtimes <- radtimes_df$datetime
my_radtimes
my_radtime <- my_radtimes[1]
my_radtime
```

> TODO: Use weather data to determine which time in L1 radiance is in L3 product

Either better weather quality product, or non-crossing flight line?

For now just use first one if there are multiple times. 

## In situ data 

Use time from L1 radiance data to get samples with minimum time difference
from in situ water chemistry and phytoplankton data

```{r}
source("R/get-closest-sample.R")
my_analytes <- c("SUVA254", "SUVA280", "doc", "tss", "chla")
my_samps <- my_analytes %>% purrr::map(~get_closest_sample(.x, my_time = my_radtime))
names(my_samps) <- my_analytes

map(my_samps, ~pull(.x, time_diff_days))
```

save the overlaps to the overlap table

```{r}
overlap_df <- map_df(my_samps, ~select(.x, namedLocation, collectDate, analyte, analyteConcentration, time_diff_days), .id = "analyteid") %>%
  distinct() %>%
  mutate(aop_siteyear = glue("{my_aq_site}{my_aop_yr}"),
        aq_site = my_aq_site, aop_yr = my_aop_yr,
        loc_type = my_loc_type,
        flightdate = as_date(my_radtime),
        notes = glue("updated {Sys.Date()}")) %>%
  dplyr::select(aop_siteyear, aq_site, aop_yr, loc_type, flightdate,
                collectDate, namedLocation, analyteid, analyte, analyteConcentration,
                time_diff_days)
  
overlap_path <- glue("results/overlaps/{my_aq_site}{my_aop_yr}.csv")
write_csv(overlap_df, overlap_path)
```

> TODO: Also get phytoplankton taxonomy for relative abundance in each group

> TODO: Figure out what to do if there are bad condition samples

> TODO: Save closest samples for each year then look for correlations
with chl a, turbidity, CDOM, DOC, at each wavelength (separate doc)

```{r}
fs::dir_ls("results/overlaps") %>%
  map_df(~read_csv(.x))
```

