---
title: "Reflectance Normalization"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(glue)
library(tidyverse)
library(readxl)
library(lubridate)
```

# Background

* Need to normalize measured (atmospherically corrected) reflectance spectra to 
**normalized water-leaving reflectance** ($[R_w]_N$) based on zenith 
illumination (angle between vertical and sun) and viewing geometry. 
* Zenith angles are in the L3 reflectance metadata. 
* `03-solar-zenith-angle.R` for extracting

see USGS reference page [here](https://www.usgs.gov/land-resources/nli/landsat/solar-illumination-and-sensor-viewing-angle-coefficient-files)

In hydrologic optics, models are sometimes defined based on zenith angle $theta$
rather than scattering angle $\psi$, for the convenience of having $\theta$ 
distinguish two hemispheres separated by the water surface. This makes parameters
AOP (apparent) rather than IOP (inherent), without a correction for the angle 
of incidence

```{r}
my_aq_site <- "BARC"
my_aop_yr <- "2017"
my_loc_type <- 'buoy.c0'
```

# Measured Reflectance Spectra

> TODO: interpolate to the same wavelengths for all years? 

```{r}
spectra_dir <- "/Volumes/hondula/DATA/spectra_buff5m/"
spectra_file <- fs::dir_ls(spectra_dir, regexp = glue("{my_aq_site}_{my_aop_yr}"))

spectra_df <- read_csv(spectra_file)
my_spectra_pt <- spectra_df %>% filter(my_cellid)

rrs <- spectra_df %>%
  ggplot(aes(wavelength, reflectance)) +
  geom_line(aes(group = cell)) +
  geom_line(data = my_spectra_pt, col = "green", lwd = 1) +
  theme_bw() +
  coord_cartesian(ylim = c(0, 0.1)) +
  theme(legend.position = "none")
rrs + ggtitle(glue("Measured reflectance at {my_aq_site} {my_loc_type} {my_aop_yr}"))

```

```{r}
my_wavelengths_rs <- my_spectra_pt$wavelength
range(my_wavelengths_rs)
```

> TODO: Maybe This is wrong? these are the sensor not solar zenith angles

Irradiance Reflectance just BELOW water surface:

$R(0 -, \lambda) = C \frac{b_b(\lambda)}{a(\lambda) + b_b(\lambda)}$

Irradiance is diffuse and radiance is directional

C is a function of solar altitude that is a linear function of $\mu_0$, the 
cosine of the zenith angle of the refracted photons (Kirk 1984)

$C(\mu_0) = -0.629\mu_0 + 0.975$

As light passes through water-air interface, it undergoes refraction that increases
its angle to the vertical. Surface factor of 0.544 (Austin 1980) relates 
radiance just above the surface with radiance just below the surface. 

**Diffuse component of remote sensing reflectance** just above the water surface:

$r_D = 0.544(-0.629\mu_0 + 0.975) \frac{b_b(\lambda)}{a(\lambda) + b_b(\lambda)}$


```{r}
# zenith_dir <- "/Volumes/hondula/DATA/zenith-angles"
# zenith_file <- fs::dir_ls(zenith_dir, regexp = glue("{my_aq_site}_{my_aop_yr}"))
# 
# zenith_df <- read_csv(zenith_file)
# my_zenith_pt <- zenith_df %>% filter(my_cellid)
# my_zenith_pt$to_sensor_zenith_angle
zenith_angle <- 32.01384 ## need to update these!!!
mu0 <- cos(zenith_angle)
mu0
C_mu0 = -0.629 * mu0 + 0.975
surface_factor = 0.544
```

Calculate normalized water leaving reflectance from surface factor and $cos(\mu_0)$

$R_{rs}$ 

$[R_w]_N = R_{rs}$

```{r}
# rD = 0.544 * C_mu0 * (backs_water/ (abs_total + backs_water)
```

